---
title: "Using Taxize"
author: "Nate Olson"
date: "May 18, 2016"
output: html_document
---
TODO - clean-up, fix samples missing classifications or taxid, add genus to org info
```{r}
library(ProjectTemplate)
load.project()
```

```{r}
single_org_uid <- singleOrgRef$uid
single_org_gi <- singleOrgRef$gi
## creating a lookup table
names(single_org_gi) <- single_org_uid
```

```{r}
library(taxize)
```

## Two Taxid Comparison Example
```{r}
taxid <- genbank2uid(id = single_org_gi[1])[[1]]
tax_class <- classification(taxid, db = "ncbi")
taxid2 <- genbank2uid(id = single_org_gi[50])[[1]]
tax_class2 <- classification(taxid2, db = "ncbi")
```

Tax Rank comparison
Need to extract last row
```{r}
inner_join(tax_class$`1415630`, tax_class2$`1048689`) %>% tail(1)
```

## Analyzing Full Single Org Dataset
### Generating Query ID Table
Table used to convert ids between uid, gi, and taxid.

* uid - used in single_org_dat table to identify query  
* gi - used to obtain taxid with `genbank2uid`  
* taxid - used to obtain classifications  

```{r}
query_id_tbl <- singleOrgRef %>% select(uid, gi)
query_gi <- query_id_tbl$gi %>% unique()

## getting taxid
query_taxid <- query_gi %>% map(~genbank2uid(id = . )) %>% map(1) %>% flatten_chr()

## joining all three

query_id_tbl <- data_frame(gi = query_gi, taxid = query_taxid) %>% 
      left_join(query_id_tbl, .)
query_uid_taxid <- query_id_tbl %>% select(uid, taxid) %>% unique()
```

### Taxonomic Classification
List of taxize taxonomic classifications for all query and match taxids
```{r}
query_taxid_classification <- query_id_tbl$taxid %>% unique() %>% 
      classification(db = 'ncbi')
```

This step is rather slow due to the api call and required timeout. 
Saving object to rds to prevent having to rerun.  
```{r}
saveRDS(query_taxid_classification, "../data/query_taxid_class.rds")
```

Retrieving match taxid classification, excluding query taxids
```{r}
match_taxid <- single_org_dat$match_taxid[
      !(single_org_dat$match_taxid %in% query_id_tbl$taxid)] %>% unique()

## error with 12908 will need to look into
bad_taxid <- c("12908","0")
match_taxid <- match_taxid[!(match_taxid %in% c(bad_taxid))]

match_taxid_classification <- list()
for(i in match_taxid){
      if(!(i %in% bad_taxid)){
          match_taxid_classification[i] <- classification(i, db = 'ncbi')    
      }
}
```

Finishing Incomplete Run
```{r}
complete_taxid <- names(match_taxid_classification)

match_taxid <- match_taxid[!(match_taxid %in% c(bad_taxid, complete_taxid))]

for(i in match_taxid){
    match_taxid_classification[i] <- classification(i, db = 'ncbi')    
}
```

```{r}
saveRDS(match_taxid_classification, "../data/single_org_match_taxid_class.rds")
```

Combined Taxonomic Classificaitons
```{r}
taxid_classification <- c(match_taxid_classification, query_taxid_classification)
```



### Appending Query IDs
Adding query ids to single org analysis results
```{r}
id_tbl <- query_uid_taxid %>% 
      mutate(uid = as.character(uid),
             taxid = as.character(taxid)) %>% 
      rename(query_uid = uid, query_taxid = taxid)
             
single_org_match_taxid <- single_org_dat %>% select(match_taxid, query_uid) %>% 
      left_join(id_tbl) %>% select(match_taxid, query_taxid) %>% unique() %>% 
      filter(!(match_taxid %in% bad_taxid),
             !is.na(query_taxid))
```

__NEED TO CHECK__ Not sure why more columns in `single_org_dat` and `single_org_match_taxid`

### Finding Lowest Common Match Level
```{r}
match_class_df <- data_frame()
for(i in 1:nrow(single_org_match_taxid)){
      match_taxid <- single_org_match_taxid$match_taxid[i]
      match_class <- taxid_classification[[match_taxid]]
      query_taxid <- single_org_match_taxid$query_taxid[i]
      query_class <- taxid_classification[[query_taxid]]
      match_class_df <- inner_join(match_class, query_class) %>% 
            filter(rank != "no rank") %>% tail(1) %>%
            mutate(query_taxid = query_taxid, match_taxid = match_taxid) %>% 
            bind_rows(match_class_df, .)
}
```

```{r}
write_csv(match_class_df,"../data/match_class_results.csv")
```

## Distribution of Taxa Level Matches
```{r}
ggplot(match_class_df) + geom_bar(aes(x = rank)) + 
      theme(axis.text.x = element_text(angle = 90))
```

## Combining Species
```{r}
match_df <- match_class_df %>% 
      mutate(rank = ifelse(rank %in% c("species", "species group",
                                       "species subgroup", "subspecies"), 
                           "species", rank))
```

## Distribution of Taxa Level Matches
```{r}
ggplot(match_df) + geom_bar(aes(x = rank)) + 
      theme(axis.text.x = element_text(angle = 90))
```

## Adding Match Results
```{r}
single_org_match_dat <- single_org_dat %>% left_join(id_tbl) %>% left_join(match_df)
```


```{r}
single_org_match_dat %>% group_by(org, rank) %>% 
      summarise(total_guess = sum(`Final Guess`)) %>% 
      mutate(rank = factor(rank, levels = c("species","genus","family","order",
                                            "class","phylum","superkingdom"))) %>% 
      ggplot() +
            geom_jitter(aes(x = rank, y = total_guess))
```

